### 2016-02-19 MSRChallenge2016.tex
Revised Sec.4

% This is "sig-alternate.tex" V2.1 April 2013
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate-05-2015}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames]{color}
\usepackage{soul} 
\usepackage{booktabs}

\newcommand{\todo}[1]
  {{\scriptsize \textbf{\color{red} {#1}}}}

\graphicspath{ {} }

\begin{document}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{10.475/123_4}

% ISBN
\isbn{123-4567-24-567/08/06}

%Conference
%\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

%\acmPrice{\$0.00}

%
% --- Author Metadata here ---
\conferenceinfo{MSR}{2016 Austin, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{A deeper look into bug fixes: Patterns, replacements, deletions, and additions}

%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Mauricio Soto\\
       \affaddr{School of Computer Science}\\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA}\\
       \email{mauriciosoto@cmu.edu}
       % 3rd. author
\alignauthor Ferdian Thung\\
       \affaddr{Singapore Management University}\\
       \affaddr{Singapore}\\
       \email{ferdiant.2013@smu.edu.sg}
% 2nd. author
\alignauthor Chu-Pan Wong   \\
       \affaddr{School of Computer Science}\\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA}\\
       \email{chupanw@cs.cmu.edu}       
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor
Claire Le Goues\\
       \affaddr{School of Computer Science}\\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA}\\
       \email{clegoues@cs.cmu.edu}
% 5th. author
\alignauthor
David Lo\\
\affaddr{Singapore Management University}\\
\affaddr{Singapore}\\
\email{davidlo@smu.edu.sg}       
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{February 2, 2016}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
  Many implementations of recent work to automatically repair bugs in software
  target programs written in C. Work that targets Java often begins from, or at
  least compares to, direct translations of such techniques to a Java context.
  However, Java and C are very different languages, and these differences should be
  studied and understood if existing approaches are to be profitably translated
  from one to another.  We conducted a large-scale study of bug-fixing commits
  to Java projects, focusing on assumptions underlying common
  search-based repair approaches.  We make a number of observations that can be
  leveraged to guide automatic software repair to target Java specifically,
  including common and uncommon statement replacements as conducted by human
  developers and the applicability of previously-proposed patch construction
  operators in the Java context.  We anticipate that these observations can
  inform more effective patch construction techniques for Java, both in raw
  numbers of bugs fixed and the usability and maintainability of the resulting
  patches. 
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

%
% End generated code
%

%
%  Use this command to print the description
%
\printccsdesc

% We no longer use \terms command
%\terms{Theory}

\keywords{Automatic error repair; Maintainability; Human-like patches}

\section{Introduction}

Research in automatic bug repair seeks to develop techniques that can
automatically construct patches to repair software bugs, which are prevalent in
practice. There has been considerable attention paid to this problem in recent
years~\cite{kim2013,weimer2009,legoues2012,pan2009,so,many,more}.
Mutation-based, or generate-and-validate approaches, typically operate by
generating a large number of candidate patches using a pre-defined set of
mutation operators, and then validating their correctness with respect to a set
of test cases.  These mutation operators range from simple, coarse granularity
statement-level mutations~\cite{examples} to human-constructed templates learned
from a large corpus of previous human bug-fixing commits~\cite{kim2013}.

A considerable proportion of this previous work is implemented to target
programs written in C, with only limited attention paid to programs in other
languages, like Java~\cite{kim2013}.  At least partially as a result,
researchers targeting Java must either start from or at least compare against
imperfect baselines, that is, Java-based implementations of techniques
originally implemented for C~\cite{nopol}. %  Even PAR~\cite{kim2013}, perhaps the
% best-known previously-proposed Java-based program repair technique, both starts
% from and is directly compared to a Java implementation of GenProg~\cite{genprog}, which
% targets the C programming language.  
This C focus extends to empirical studies
of bug fixing practice, used to inform research questions in automatic
repair~\cite{zhong2015}.

This is an oversight, because, to put it simply: Java and C are very different
languages.  Consider GenProg~\cite{genprog,cites}, a well-known technique in
this space: GenProg uses a genetic programming heuristic to combine
statement-level mutations into patches to address a particular defect.
``Statement'' is a well-defined semantic unit in the C programming
language~\cite{cspec}; to which semantic unit in Java does this concept map?
Can we learn more
intelligent rules to inform candidate modifications
(cf. Prophet~\cite{rinardPopl2016}, for C)?  What types of
semantic scoping rules should be used to limit mutations, given the relative
stringence of the Java compilers?

In this paper, we study bug-fixing commits to Java programs, taking as a subject
of study several million human-made bug fixes from Github. We study broad
characteristics of changes to fix bugs; the applicability of previously-proposed
templates to repairing defects in Java programs~\cite{kim2013}; and the nature
of additions, replacements, of statements in bug-fixing commits. We make
observations to directly guide future research in automatic repair to Java
programs, to increase the success rate of such techniques and the degree to
which the patches are human-like and therefore more readable and maintainable by
human developers; this is important, because patch quality is an important
concern in this research area~\cite{Qi2013}.

% On the other hand, some approaches~\cite{kim2013} search for specific patterns
% and modify source code according to predefined templates. A well known approach
% is PAR~\cite{kim2013}, which creates 10 different repair templates and applies
% them to the buggy code in an effort to repair it. In this paper we have taken 8
% out of the 10 PAR templates and tested how common they are in the repairs made
% by programmers in the latest official data dump of Github as of September 2015
% provided by~\cite{dyer2013}.  Our results provide evidence of how common those
% patterns are in practice. We take as reference the study performed by Dongsun
% Kim et al.~\cite{kim2013} in which they look for the most common ways in which
% programmers patch bugs in software. The researchers developed a variation of the
% tool Genprog~\cite{weimer2009,legoues2012} with several different templates
% resembling patterns programmers use to patch bugs, and then performed an
% empirical study to evaluate which patches do human programmers prefer.

A good complement for this paper is~\cite{zhong2015}, where they look for
additions and deletions in a smaller data set. In this paper, we focus more on
the replacements, which have been left out of the analysis of~\cite{zhong2015},
and expand their results of the deletions and additions in a broader
dataset~\cite{dyer2013}.
\todo{note to self (CLG) should mention the work from Earl, Yuriy, and others
  that looked at the redundancy assumption.}

\todo{roadmap}

\section{Dataset and Characteristics}

We begin by broadly characterizing the bug-fixing changes to Java code in our
dataset.  We perform our analysis on the \emph{2015 September/Github} dataset on
Boa, which includes $7,830,023$ Java projects with $23,229,406$ revisions. Among
all the revisions, $4,590,679$ revisions are identified by Boa as bug fixing
revisions.

\subsubsection*{How many files are changed to fix a bug?} \vspace{-0.3em}

A file was changed if (1) it was a new file; (2) or it was modified; (3) or it
was deleted. In total, $52,052,571$ files got changed. Thus, on average each bug
fixing revision changes $11.3$ files.  This number is surprisingly high because
most automatic program repair techniques assume that bugs are local in most
cases. There could be several reasons behind this number. It could be the case
that other software artifacts (e.g.  documentation) are updated after bug
fixing. It could also be the case that bug fixing commit also contains changes
that are not related to fixing bug (e.g. refactoring, adding new features by
creating new files). To better understand why so many files are changed, we
continue to ask:

\subsubsection*{What are the file types of those changed files?} \vspace{-0.3em}

To answer this question, we iterate all the changed files in each bug fixing
revision and count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.

\begin{table}
\centering
  \begin{tabular}{| c | r | r |}
  \hline
  File Kind & Total Number & Average \\ \hline \hline
  BINARY & 752,945 & 0.16 \\ \hline
  SOURCE\_JAVA\_ERROR & 2,073,558 & 0.45 \\ \hline
  SOURCE\_JAVA\_JLS2 & 2,607,413 & 0.57 \\ \hline
  SOURCE\_JAVA\_JLS3 & 15,748,967 & 3.43 \\ \hline
  SOURCE\_JAVA\_JLS4 & 83,798 & 0.02 \\ \hline
  TEXT & 541,023 & 0.12 \\ \hline
  XML & 6,818,299 & 1.49 \\ \hline
  UNKNOWN & 23,426,568 & 5.10 \\ \hline
  \end{tabular}
  \caption{File Types of Changed Files}
  \label{tbl:fileType}
\end{table}

SOURCE\_JAVA\_ERROR file kind represents files that Boa failed to parse. JLS2,
JLS3, and JLS4 represent Java files written in different J2SE version. Boa
documentation does not explicitly explain what is TEXT and what is UNKNOWN. From
Table~\ref{tbl:fileType}, we can see that TEXT files and BINARY files are
changed least frequently. This is not surprising because TEXT files are usually
documentation files, and BINARY files should only be changed in rare case (e.g.
milestone, new release, dependency update). If we consider all 4 kinds of Java
files as one kind, then on average each bug fixing revision changes $4.47$ Java
files. But still, it does not conform to our assumption of bug locality. XML
files in Java projects usually represents build files, thus changing one build
file after each bug fixing is not surprising.  What surprises us is that UNKNOWN
files are changed most frequently. Note that all the projects in our dataset are
Java projects, so it is reasonable to assume that these files are not source
files that written in other programming languages.  One possible explanation is
that they are test resources. However, due to the limitations of Boa, we can not
look into the content of these files.

It could be the case that among those $4.47$ changed Java files, some of them
are related to testing, such as modifying the test cases and creating new test
classes. To better understand what changes are usually made to Java source code
files, we continue to ask:

\subsubsection*{How often do developers introduce new classes, methods, fields,
	local variables in
bug fixing commits?} \vspace{-0.3em}

This question is interesting because most automatic program repair approaches
often do not create new classes/methods/fields/local variables to fix bugs.
Most approaches guess and synthesize the patches based on existing source code,
and direct translations of approaches from C to Java do not naturally consider
object-oriented concerns, like fields or classes.

Table~\ref{tbl:new} shows our analysis results. It shows that introducing new
classes and new variables are rare, which is good news for automatic program
repair approaches. On average, we still have $0.69$ new method per bug fix. If
we consider the possibility of creating new test method (i.e.  methods that
annotated by @Test if JUnit is used), this number is not that scary. However,
what surprises us is that $1.32$ fields are created per bug fix.  This indicates
that automatic program repair approaches should pay attention to the state of
the class when fixing Java programs.

\begin{table}
\centering
  \begin{tabular}{| c | r | r |}
  \hline
  Introducing & Total Number & Average \\ \hline \hline
  Class & 729,201 & 0.16 \\ \hline
  Methods & 3,186,867 & 0.69 \\ \hline
  Fields & 6,076,646 & 1.32 \\ \hline
  Variables & 924,259 & 0.2 \\ \hline
  \end{tabular}
  \caption{Introducing new classes, methods, fields, local variables to fix bugs}
  \label{tbl:new}
\end{table}

% \todo{Is there room/call to include any element of Chu Pan's replacement
%   analysis in here?}

\section{PAR Bug Fixing Patterns}\label{sec:method}

Par is a well-known technique for patch generation that 
On the other hand, some approaches~\cite{kim2013} searches for specific patterns
and modifies java code according to predefined templates, constructed by humans
to cover a large set of bug fixes from an existing corpus.  As one of the only
techniques to directly target Java, it is an important source of possible
mutations to study to inform future directions. 
We therefore search for and analyze most of the bug fixing patterns proposed by
Kim et al.~\cite{kim2013} and assess their prevalence on this distinct
dataset by using Boa's language infrastructure~\cite{dyer2013} to identify the patterns
in the commits in the dataset.

% A well known approach
% is PAR~\cite{kim2013}, which creates 10 different repair templates and applies
% them to the buggy code in an effort to repair it. In this paper we have taken 8
% out of the 10 PAR templates and tested how common they are in the repairs made
% by programmers in the latest official data dump of Github as of September 2015
% provided by~\cite{dyer2013}.  Our results provide evidence of how common those
% patterns are in practice. We take as reference the study performed by Dongsun
% Kim et al.~\cite{kim2013} in which they look for the most common ways in which
% programmers patch bugs in software. The researchers developed a variation of the
% tool Genprog~\cite{weimer2009,legoues2012} with several different templates
% resembling patterns programmers use to patch bugs, and then performed an
% empirical study to evaluate which patches do human programmers prefer.

\subsection{Detecting PAR patterns in Boa}

Boa's capabilities are powerful, but are limited in the precision it enables in
our detection of these patterns. For example, it is unable to directly perform a
diff between two file versions. Thus, rather than finding the exact count of
bug fixing patterns, we will find an approximate count by processing pre- and
post- fix files separately.  Additionally, there are
two of the originally proposed 10 patterns that cannot be easily detected by
Boa, as we describe below.

The patterns we search for are as follows, including our mechanism for
detecting them using Boa:

\vspace{1ex}
\noindent {\bf Altering method parameters (AMP)}.  This pattern changes the
input for method parameters.  {\bf Example:} obj.method(v1,v2) $\rightarrow$
obj.method(v1,v3). To detect this pattern, for both pre-fix and post-fix version
of a buggy file, we create a custom method call signature. This signature
contains method name, literal parameter, and variable parameter. we consider
parameter containing custom expression as OTHER. This is due to inexistence of
function that can print Abstract Syntax Tree (AST) back to source code in
Boa. We discard method signatures that appear both in pre-fix and post-fix
version. We then find whether there exists method with the same name and the
same number of parameters, but different parameter signatures in pre-fix and
post-fix version. If it exists, we consider that AMP pattern is found.

\vspace{1ex}
\noindent {\bf Calling another method with the same parameters (MSM)}.  This
pattern changes a method name.  {\bf Example:} obj.method1(param) $\rightarrow$
obj.method2(param).  To detect this pattern, we create method signature similar
like when detecting AMP pattern. This time, we find whether there exists method
with the exact same parameter signature but different name in pre-fix and
post-fix version. If it exists, we consider that MSM pattern is found.

\vspace{1ex}
\noindent
{\bf Calling another overloaded method with one more parameter (COM)}. This
pattern adds one more parameter to a method. 
	{\bf Example:} obj.method(v1) $\rightarrow$ obj.method(v1,v2).
  To detect COM pattern, we create method signature similar like when detecting
  AMP and COM pattern. This time, we find whether there exists method with
  the same name, but the number of parameters in pre-fix and post-fix version
  differs by one. If it exists, we consider that COM pattern is found.

\vspace{1ex}
\noindent
 {\bf Changing a branch condition (CBC)}. This pattern adds or removes a condition.
	{\bf Example:} if(a != b) $\rightarrow$ if(a != b \&\& c == 0).
  To detect CBC pattern, we count the number of { \em logical and} and {\em
    logical or} inside if conditional expression for both pre-fix and post-fix
  version. we assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there is a larger count in post-fix version, we consider that CBC pattern is found.

\vspace{1ex}
\noindent
 {\bf Initializing an object (IAO)}. This pattern adds a initialization to object declaration.
  To detect IAO pattern, we count the number of NEW expression in variable
  declaration for both pre-fix and post-fix version. If there is a larger count in post-fix version, we consider that IAO pattern
  is found.
	{\bf Example:} Type obj $\rightarrow$ Type obj = new Type().

\vspace{1ex}
\noindent
 {\bf Adding a null checker (ANC)}.  This pattern inserts a condition to check whether an
    object is null or not in order to prevent unexpected state.
	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}
  To detect ANC pattern, we count the number of if conditional expression that
  contains {\em !=null} or {\em ==null} for both pre-fix and post-fix
  version. If there is a larger count in post-fix version,
  we consider that ANC pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding an array out of bound checker (AOB)}. This pattern inserts a condition to check that an array
    index is within bound right before the index is used.
	{\bf Example:} arr[idx]=0 $\rightarrow$ if(idx<arr.length){arr[idx]=0}.
  To detect AOB pattern, we count the number of if conditional expression that
  contains {\em expr<var.length} or {\em var.length>expr} for both pre-fix and
  post-fix version. If there is a larger count in post-fix version, we consider that AOB pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a collection out of bound checker (COB)}.  This pattern inserts a condition to check that a
    collection index is within bound before it is used.
	{\bf Example:} x=col.get(idx) $\rightarrow$ if(idx<col.size()){x=col.get(idx)}\\
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre-fix and
post-fix version. If there are count differences between pre-fix and post-fix
version, we consider that COB pattern is found.

\vspace{1ex} The above patterns only cover 8 out of 10 bug fixing patterns in
PAR. The other two patterns are: (1) checking that an object is an instance of a
target class before casting it; and (2) replacing expression in an if condition
by another expression in the same scope. Pattern (1) cannot be analyzed because
Boa does not support {\em instanceof} expression. Similarly, pattern (2)
requires us to track scope between pre-fix and post-fix version, which is not
supported by Boa. Implementing them ourselves might be possible, but it might
require too much effort than necessary. We speculate that investigating 8 out of
10 patterns should give us a good idea on how common the patterns actually are.

\subsection{Results} \label{sec:freqfixpattern}

\begin{table}[!htb]
	\centering
	\begin{tabular}{lrr} 
		\hline
		& \textbf{\#Appearance} & \textbf{Percentage}\\
		\hline
		%Buggy Files & 46,301,429 \\ 
		AMP Pattern & 901,083 & 1.95\%\\
		MSM Pattern & 783,073 & 1.69\%\\
		COM Pattern & 270,128 & 0.58\%\\ 
		CBC Pattern & 1,959,377 & 4.23\%\\  
		IAO Pattern & 1,308,006 & 2.82\%\\  
		ANC Pattern & 1,340,561 & 2.90\%\\  
		AOB Pattern & 128,016 & 0.28\%\\  
		COB Pattern & 262,915 & 0.57\%\\   
		\hline
	\end{tabular}
	\caption{Frequency of Bug Fixing Patterns}\label{tab:freqpattern}

\end{table}

\begin{table*}
	\centering
	\resizebox{\textwidth}{100pt}{
		\begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|} \hline
			&Assert&Break&Continue&Do& For&If&Label&Return&Case&Switch& Synch&Throw&Try&TypeDecl&While
			\\ \hline

\texttt{Assert}	&-&7.48&3.76&0.53&8.30&23.05&0.31&20.04&4.90&4.62&1.30&13.50&7.23&0.03&4.95
\\ \hline
\texttt{Break} &1.00&-&4.08&0.60&9.93&26.03&0.13&25.39&2.48&1.57&1.79&8.39&11.73&0.10&6.77
\\ \hline
\texttt{Continue} &1.74&9.42&-&1.28&11.39&18.25&0.35&22.60&3.80&2.85&2.17&8.98&9.42&0.11&7.63
\\   \hline
\texttt{Do} &0.81&5.26&6.60&0.00&9.44&14.21&0.18&15.86&3.73&1.67&1.97&5.88&6.39&0.03&27.98
\\ \hline
\texttt{For} &0.86&6.28&3.19&0.79&-&22.89&0.09&21.08&5.01&3.34&1.87&10.01&10.71&0.08&13.79
\\ \hline
\texttt{If}&1.64&8.43&2.87&0.60&13.49&-&0.24&26.46&7.45&4.80&2.85&9.89&15.11&0.08&6.11
\\ \hline
\texttt{Label} &1.30&8.33&7.86&1.11&5.18&22.85&-&15.17&3.05&2.04&14.62&10.45&4.16&0.09&3.79
\\ \hline
\texttt{Return} &1.13&9.41&3.11&0.49&13.33&27.24&0.24&-&5.59&3.65&2.55&14.91&12.61&0.12&5.61
\\ \hline
\texttt{Case} &0.78&2.84&2.84&0.39&10.27&31.79&0.16&22.40&-&0.46&2.07&7.37&11.69&0.08&6.87
\\ \hline
\texttt{Switch}&1.14&2.72&3.80&0.55&11.07&34.14&0.13&21.86&0.75&-&1.53&8.65&9.02&0.05&4.58
\\ \hline
\texttt{Synch} &0.80&6.57&2.28&0.43&10.21&24.18&0.05&19.77&6.35&2.07&-&9.16&12.16&0.04&5.93
\\ \hline
\texttt{Throw} &2.11&6.57&2.58&0.48&11.87&18.84&0.17&32.28&4.64&3.30&2.74&-&10.08&0.07&4.27
\\ \hline
\texttt{Try}&0.71&7.41&3.02&0.66&11.73&27.75&0.11&23.24&5.63&2.65&2.58&8.99&-&0.09&5.42
\\ \hline
\texttt{TypeDecl} &0.00&4.51&7.52&1.00&10.28&21.05&0.50&17.79&6.02&1.75&2.01&9.27&11.53&-&6.77
\\ \hline
\texttt{While}& 0.72&8.02&3.82&1.96&23.16&19.78&0.12&16.48&6.56&3.09&1.64&6.81&7.80&0.04&-
			\\ \hline
		\end{tabular}
		}
		\caption{Likeliness to replace the statement in the first column for the statement in the top row, in Java}\label{tab:likeliness}
\end{table*}

After running the bug pattern detection approach on Boa's dataset, we obtained
the result shown in Table~\ref{tab:freqpattern}. It shows that the
most common pattern among these patterns is CBC and the least common pattern is
AOB. If we assume that these patterns never appear together,
they cover 14.78\% of buggy files. Kim et al. mentioned that the
patterns that they use cover almost 30\% of real
patches~\cite{kim2013}. Although we only detect 8 out of 10 bug fixing patterns,
it is hard to believe that the remaining two patterns cover 15\% of buggy files,
especially considering that 14.78\% is a number that we get when we assume that the patterns never appear together. Moreover, the percentage of each pattern is based
on imprecise approximation which is likely to represent an upper bound rather
than a lower bound. This suggests that common bug fixing patterns might not be
as common as it appeared to be.

\section{Statement-level mutations}
Some automatic program repair approaches mutate source code using predefined
mutation operators such as addition, deletion and replacement. In
order to generate patches that make more sense to human developers, it would be
beneficial to analyze how developers mutate source code to fix bugs.

\subsection{Replacements}

Because direct diffs are difficult to identify, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another across
this dataset.  To do this, for each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.  For each statement type (as tagged by the Boa infrastructure
for Java), we process the results as follows: if number of occurrences of one
statement type decreased post-fix and the number of another type increased, we
say that the first statement type was replaced by the second statement type for
that file. It is important to notice that this analysis doesn't distinguish the
replacement of the same statement kind, since we are counting the amount of
appearances of each statement kind.

The numbers in the slots of Table~\ref{tab:likeliness} \todo{shall we add \% in
this table?}, represent the percentage rate at which the elements of the
leftmost column were replaced by the elements of the top row. In the first
column we have the list of statements that are being replaced (faulty statements
that were removed), and on the top row we have the list of statements for which
that statement was replaced for (the new statement put instead). For example, if
we take the intersection of row 5 (DoStatement row) and column 2 (AsserStatement
column) of the data, we have a 0.81, which means that Do statements were
replaced by Assert statements 0.81\% of the times.

Our results show that the most common replacement is replacing Return statement
with If statement, which showed up in 30,489 files. The second most common
replacement is replacing an If statement with a Return statement, which had
28,536 appearances. We found that the most common \emph{Replacer} statement kind
(the statement kind that most commonly replaces others) is the \textbf{If
Statement} with 101,366 appearances; and the least common \emph{Replacer} is the
\textbf{Type Declaration Statement} with 447 appearances.

On the opposite side of the data, the least common replacement was an Assert
statement replacing a Type Declaration statement, which had zero appearances.
The second least common replacement were replacing a Do statement with a Type
Declaration statement and replacing a Label with a Type Declaration, both of
which had one appearance respectively in our search. The most common $Replacee$
(the most likely statement kind to be replaced by another statement kind) is the
\textbf{Return Statement} with 111,938 appearances; and the least common
$Replacee$ is the \textbf{Type Declaration Statement} with 399 appearances.

\subsection{Deletions and Insertions}

\begin{figure*}[!t] \centering \includegraphics[height=0.3\textheight,
	width=0.8\textwidth]{StmtType} \caption{For each kind of statement, how many
	bug fixing revisions add/delete statements of the same kind?}
	\label{fig:StmtType} \end{figure*}

For each bug fixing revision $r$ and each statement kind $k$, we count how many
statements of kind $k$ are there in revision $r$ and $r-1$ ($r-1$ means the
revision before revision $r$), and then compare the two numbers.

From Figure~\ref{fig:StmtType}, we can see that, expression, if, return, for and
try catch statements are added most often. Meanwhile, the same types of
statements are also deleted most often. This finding indicates that most bugs
were fixed by changing the control flow of the program.

\subsection{Applicability to Automatic Error Repair}

This study is directly applicable to automatic software repair. Most common
tools for automatic software repair~\cite{kim2013,weimer2009,legoues2012} first
find a faulty statement using several different fault localization
techniques~\cite{fry2010}, and once they have that faulty statement, they apply
different kinds of edits to it, usually at random. Which makes it equally likely
to replace an statement for another without looking at the kinds of statements
being replaced. With this study, we are able to make a more accurate guess based
on the phenomenon we observed from Table~\ref{tab:likeliness} and
Figure~\ref{fig:StmtType}.

For example, if we are fixing a Java program, and we have a faulty statement
that happens to be a For Statement. Then we can look for the most likely
statements to replace a For Statement (row 6): If statement (22.89\%
likeliness), Return statement (21.08\%), and a While statement (13.79\% of
likeliness to replace a For Statement). We also know that there is a 7.9\% of
cases in which human programmers deleted this kind of statement, and in 10.6\%
of the cases, a For statement was added to patch a bug. This way we provide
guidelines to the automatic bug repair approaches to consider this on their
effort to find a patch.

\section{Threats to Validity}

Since we are using a DSL provided by Boa to implement our analysis, the
correctness of our analysis depends on our programs as well as Boa itself. To
prevent errors from our side, we open source our implementation on
Github~\footnote{\url{https://github.com/}}.\todo{update
this}

Some threats to our study are due to the limitation of Boa. For example, we rely
on Boa to identify bug fixing revisions, but it is unclear whether those
identified revisions are actually related to bug fixing. After all, identifying
bug fixing revisions from revision history is still an open problem. Another big
limitation of Boa is that there is no easy way to get a precise statement-level
diff between two revisions. Thus, our template matching and analysis of code
changes (by counting each statement kind) could be inaccurate. However, our
results should provide upper bounds for the questions we investigated.

\section{Conclusions}
We have characterized bug fixing changes in Java.
On average, a bug fixing revision changes 11.3 files and often introduces new
fields and methods. We have also showed that 8 out of 10 bug fixing patterns in
PAR cover only 15\% of bug fixes. Last but not least, we investigated statement
level mutations and summarized the likelihood of statement insertion, deletion,
and replacement.

The findings of our study provide a set of useful insights for automatic
program repair tools in Java. It suggests that patterns in PAR are not enough to
cover the entirety of bug fixes in Java. Adding more patterns is one solution,
but may not be a permanent one. Our finding on Java bug fixes characteristic suggests that mutation based program repair may need to add new fields or methods
to fix bugs in Java and achieve a human-like written patch. A deeper look suggests
that it can also improve its search for a patch by having the knowledge of how
likely it is to replace a faulty statement by the different kinds of statements
available. Overall, our findings demonstrate that our knowledge of bug fixing
properties in Java is still lacking and therefore more research on it are
warranted.

%\end{document}  % This is where a 'short' article might terminate

\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns

%\balancecolumns % GM June 2007
% That's all folks!
\end{document}