### 2016-02-19 MSRChallenge2016.tex
half a column over

% This is "sig-alternate.tex" V2.1 April 2013
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate-05-2015}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames]{color}
\usepackage{soul} 
\usepackage{booktabs}

\newcommand{\todo}[1]
  {{\scriptsize \textbf{\color{red} {#1}}}}

\graphicspath{ {} }

\begin{document}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{10.475/123_4}

% ISBN
\isbn{123-4567-24-567/08/06}

%Conference
%\conferenceinfo{PLDI '13}{June 16--19, 2013, Seattle, WA, USA}

%\acmPrice{\$0.00}

%
% --- Author Metadata here ---
\conferenceinfo{MSR}{2016 Austin, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{A deeper look into bug fixes: Patterns, replacements, deletions, and additions}

%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Mauricio Soto$^1$, Ferdian Thung$^2$, Chu-Pan Wong$^1$, Claire Le Goues$^1$, and David Lo$^2$\\
       \affaddr{$^1$School of Computer Science, Carnegie Mellon University, Pittsburgh, PA\\
       $^2$Singapore Management University, Singapore}\\
       \email{mauriciosoto@cmu.edu, \{ferdiant.2013,davidlo\}@smu.edu.sg, \{chupanw,clegoues\}@cs.cmu.edu} 
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{February 2, 2016}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
  Many implementations of research techniques that automatically repair software
  bugs target programs written in C. Work that targets Java often begins from, or at
  least compares to, direct translations of such techniques to a Java context.
  However, Java and C are very different languages, and these differences should be
  studied and understood if existing approaches are to be profitably translated
  between them.  We conducted a large-scale study of bug-fixing commits
  to Java projects, focusing on assumptions underlying common
  search-based repair approaches.  We make observations that can be
  leveraged to guide high quality automatic software repair to target Java specifically,
  including common and uncommon statement replacements as conducted by human
  developers and the applicability of previously-proposed patch construction
  operators in the Java context. 
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

%
% End generated code
%

%
%  Use this command to print the description
%
\printccsdesc

% We no longer use \terms command
%\terms{Theory}

\keywords{Automatic error repair; Maintainability; Human-like patches}

\section{Introduction}

There has been considerable attention paid to the problem of automatic
program repair in recent
years~\cite{kim2013,legoues2012,pan2009,
  Mechtaev15,prophet}, among many others),
%
One broad class of techniques in this space take a generate-and-validate
approach, which involves generating a large number of candidate patches using a
pre-defined set of mutation operators, and then validating correctness with
respect to a set of test cases.  Mutation operators range from simple, coarse
granularity statement-level mutations to human-constructed templates learned
from a large corpus of previous human bug-fixing commits.

A considerable proportion of this works targets programs written in C.
Researchers targeting Java often start from or compare against imperfect
baselines, that is, Java-based implementations of techniques originally
implemented for C~\cite{nopol,kim2013}. % Even PAR~\cite{kim2013}, perhaps the
% best-known previously-proposed Java-based program repair technique, both starts
% from and is directly compared to a Java implementation of GenProg~\cite{genprog}, which
% targets the C programming language.  
%
This is an oversight, because, frankly Java and C are very different languages.
Consider GenProg~\cite{legoues2012}, which
combines statement-level mutations into patches to address a particular
defect. Which ``Statements'' (a semantic unit in C) should be manipulated in Java?
Can we learn
intelligent rules to inform candidate modifications
(cf. Prophet~\cite{Long2016}, for C)?  What types of semantic scoping
should limit mutations, given the relative stringence of the Java
compiler?

In this paper, we study bug-fixing commits to Java programs, taking as a subject
of study several million human-made bug fixes from Github. We study broad
characteristics of these changes; the applicability of previously-proposed
Java repair templates~\cite{kim2013}; and the nature
of additions and replacements of Java statements in bug-fixing commits. We make
observations to directly guide future research in automatic repair of Java
programs, to increase the success rate of such techniques and the degree to
which the patches are human-like and therefore more readable and maintainable by
human developers. This is important, because patch quality is an important
concern in this research area~\cite{Qi15}.

% On the other hand, some approaches~\cite{kim2013} search for specific patterns
% and modify source code according to predefined templates. A well known approach
% is PAR~\cite{kim2013}, which creates 10 different repair templates and applies
% them to the buggy code in an effort to repair it. In this paper we have taken 8
% out of the 10 PAR templates and tested how common they are in the repairs made
% by programmers in the latest official data dump of Github as of September 2015
% provided by~\cite{dyer2013}.  Our results provide evidence of how common those
% patterns are in practice. We take as reference the study performed by Dongsun
% Kim et al.~\cite{kim2013} in which they look for the most common ways in which
% programmers patch bugs in software. The researchers developed a variation of the
% tool Genprog~\cite{weimer2009,legoues2012} with several different templates
% resembling patterns programmers use to patch bugs, and then performed an
% empirical study to evaluate which patches do human programmers prefer.

\noindent\textbf{Related work.} Zhong and Su~\cite{zhong2015} study
a similar set of questions in C.  Beyond focusing on Java, we focus more on
replacements, ommitted from that previous analysis, and target a
larger dataset~\cite{dyer2013}.  Barr et al.~\cite{Barr14fse} studied changes to
Java programs to understand the ``Plastic Surgery Hypothesis'' underlying
certain types of program repair, an important but orthogonal
concern. Kim et al.~\cite{kim2013} manually analyze 
changes to Java programs to inform an automated repair
technique and show that doing so results in higher-quality repairs; 
we attempt to validate their templates heuristically on
a different dataset, and take their results as
motivation for understanding human repairs (as it may result in better patches).
Long and Rinard learn probabilistic models from bug fixes to C
code~\cite{Long2016}. Our analysis is significantly less precise and does not
inform a new repair technique.  Rather, it serves as a starting
point for research investigating the repair of bugs in Java code.

\section{Dataset and Characteristics}

We begin by broadly characterizing the bug-fixing changes to Java code in the
dataset.  We perform our analysis on the \emph{2015 September/Github} dataset on
Boa, including $7,830,023$ Java projects with $23,229,406$ revisions. Boa
identifies $4,590,679$ as bug fixing.

\vspace{1ex}
\noindent\textbf{How many files are changed to fix a bug?}
%
We consider a file changed if it is new, modified, or
deleted in a commit. In total, $52,052,571$ files were changed. Thus, on average each bug
fixing revision changes $11.3$ files.  This is surprisingly high:
most automatic program repair techniques assume that bugs are local.
We hypothesize several explanations. Other
software artifacts (e.g.  documentation) may be updated after bug  
fixing, or fixing commits may contain changes
unrelated to fixing bug (e.g. refactoring, adding new features by
creating new files). To better understand this
phenomenon, we ask:

\vspace{1ex}
\noindent\textbf{What are the file types of those changed files?}
%
To answer this question, we iterate over all changed files in each bug fixing
revision and count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.

\begin{table}
\centering
  \begin{tabular}{ l | r  r }
  \toprule
  File Kind & Total & Average \\ 
  \midrule
  Binary & 752,945 & 0.16 \\ 
  Java (ERROR) & 2,073,558 & 0.45 \\ 
  Java (JLS2) & 2,607,413 & 0.57 \\ 
  Java (JLS3) & 15,748,967 & 3.43 \\
  Java (JLS4)  & 83,798 & 0.02 \\ 
  Text & 541,023 & 0.12 \\ 
  XML & 6,818,299 & 1.49 \\ 
  UNKNOWN & 23,426,568 & 5.10 \\ 
\bottomrule
  \end{tabular}
  \caption{File Types of Changed Files}
  \label{tbl:fileType}
\end{table}

JAVA (ERROR) files are those that Boa failed to parse. JLS2,
JLS3, and JLS4 represent Java files written in different J2SE versions. From
Table~\ref{tbl:fileType}, we see that Text and Binary files are
changed least frequently. This is unsurprising, since text are often
documentation files, and binaries should only be changed rarely. If we consider all 4 kinds of Java
files as one kind, on average, each bug fixing revision changes $4.47$ Java
files. This still does not conform to our assumption of bug locality. XML
files in Java projects usually represents build files; changing a build
file is unsurprising.  Rather more surprising is how frequently UNKNOWN files
are changed. Given that all the projects are Java projects, 
it is unclear whether these are written in other proramming
languages; One possibility is
that they are test resources. However, due to the limitations of Boa, we can not
look into the content of these files. 
\todo{Possibly mention median.} 
To better understand what changes are usually made to Java source code
files, we next ask:

\vspace{1ex}
\noindent\textbf{How often do developers introduce new classes, methods, fields,
	local variables in
bug fixing commits?} 
%
This question is interesting because most automatic program repair approaches
often do not create new methods or variables to fix bugs in C.
Most approaches guess and synthesize the patches based on existing source code,
and direct translations of approaches from C to Java do not naturally consider
object-oriented concerns, like fields or classes.

Table~\ref{tbl:new} shows that introducing new
classes and new variables is rare. On average, we still have $0.69$ new methods
per bug fix. These methods naturally contain new test methods (annotated by
@Test, for example), which is not too alarming. However,
what surprises us is that $1.32$ fields are created per bug fix.  This indicates
that automatic program repair approaches should pay attention to the state of
the class when fixing Java programs.

\begin{table}
\centering
  \begin{tabular}{l | r  r }
  \toprule
  Introducing & Total Number & Average \\ \midrule
  Class & 729,201 & 0.16 \\ 
  Methods & 3,186,867 & 0.69 \\ 
  Fields & 6,076,646 & 1.32 \\ 
  Variables & 924,259 & 0.2 \\ 
\bottomrule
  \end{tabular}
  \caption{Introducing new classes, methods, fields, local variables to fix bugs}
  \label{tbl:new}
\end{table}

% \todo{Is there room/call to include any element of Chu Pan's replacement
%   analysis in here?}

\section{Bug Fixing Patterns}\label{sec:method}

One of the best-known program repair techniques to target Java is
PAR~\cite{kim2013}.  Par searches for specific patterns
and modifies java code according to predefined templates, constructed by humans
to cover a large set of bug fixes from an existing corpus.  Par serves as an
important source of possible 
mutations to study to inform future directions. 
We therefore search for and analyze most of the bug fixing patterns proposed by
Kim et al. and estimate their prevalence in this
dataset.

% A well known approach
% is PAR~\cite{kim2013}, which creates 10 different repair templates and applies
% them to the buggy code in an effort to repair it. In this paper we have taken 8
% out of the 10 PAR templates and tested how common they are in the repairs made
% by programmers in the latest official data dump of Github as of September 2015
% provided by~\cite{dyer2013}.  Our results provide evidence of how common those
% patterns are in practice. We take as reference the study performed by Dongsun
% Kim et al.~\cite{kim2013} in which they look for the most common ways in which
% programmers patch bugs in software. The researchers developed a variation of the
% tool Genprog~\cite{weimer2009,legoues2012} with several different templates
% resembling patterns programmers use to patch bugs, and then performed an
% empirical study to evaluate which patches do human programmers prefer.

\subsection{Detecting PAR patterns in Boa}

Boa's capabilities are powerful, butlimited in the precision it enables in
our detection of Par bug-fixing patterns. For example, it is unable to directly perform a
diff between two file versions. Thus, rather than finding the exact count of
bug fixing patterns, we find an approximate count by processing pre- and
post- fix files separately.  Additionally, there are
two of the originally proposed 10 patterns that cannot be easily detected by
Boa, as we describe below. 
%
The patterns we search for are as follows, including our mechanism for
detecting them using Boa:

\vspace{1ex}
\noindent {\bf Altering method parameters (AMP)}.  This pattern changes 
input method parameters. %(\texttt{obj.method(v1,v2) $\rightarrow$
%obj.method(v1,v3)}). 
To detect this pattern, for both pre-fix and post-fix version
of a buggy file, we create a custom method call signature that
contains the name, literal parameters, and variable parameters (more complex
expressions are listed as OTHER).
%  This is due to inexistence of
% function that can print Abstract Syntax Tree (AST) back to source code in
% Boa. 
We discard method signatures that appear both in pre-fix and post-fix
versions, and then identify methods with the same name and the
same number of parameters, but different parameter signatures in pre- and
post-fix versions.% If it exists, we consider that AMP pattern is found.

\vspace{1ex}
\noindent {\bf Calling another method with the same parameters (MSM)}.  This
pattern changes a method name. % (\texttt{obj.method1(param) $\rightarrow$
%obj.method2(param)}).  
To detect this pattern, we create method signature similar
to the above for the AMP pattern, looking for methods
with the exact same parameter signature but different names in pre- and
post-fix versions. %If it exists, we consider that MSM pattern is found.

\vspace{1ex}
\noindent
{\bf Calling another overloaded method with one more parameter (COM)}. This
pattern adds one more parameter to a method. %(\texttt{obj.method(v1) $\rightarrow$ obj.method(v1,v2)}).
We create method signature similar to the above, seeking methods with
  the same name, one more parameter in the post-fix version.

\vspace{1ex}
\noindent
 {\bf Changing a branch condition (CBC)}. This pattern adds or removes a condition.
%(e.g., \texttt{if(a != b) $\rightarrow$ if(a != b \&\& c == 0)}).
  To detect CBC pattern, we count the number of { \em logical and} and {\em
    logical or} inside if conditional expression sfor both pre-fix and post-fix
  version. We assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there is a larger count in post-fix version,
  we consider the CBC pattern is found.

\vspace{1ex}
\noindent
{\bf Initializing an object (IAO)}. This pattern adds a initialization to object
declaration.  We count the number of NEW expressions in variable declarations
for both pre- and post-fix versions. If there is a larger count in post-fix
version, we consider that IAO pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a null checker (ANC)}.  This pattern inserts a condition to check whether an
    object is null.
%	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}
  To detect ANC pattern, we count the number of if conditional expressions that
  contains {\em !=null} or {\em ==null} for both pre-fix and post-fix
  version. If there is a larger count in post-fix version,
  we consider that ANC pattern found.

\vspace{1ex}
\noindent
 {\bf Adding an array out of bound checker (AOB)}. This pattern inserts a condition to check that an array
    index is within bound right before the index is used.
%	{\bf Example:} arr[idx]=0 $\rightarrow$ if(idx<arr.length){arr[idx]=0}.
 We count the number of if conditional expression that
  contains {\em expr<var.length} or {\em var.length>expr} for both pre-fix and
  post-fix version. If there is a larger count in post-fix version, we consider that AOB pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a collection out of bound checker (COB)}.  This pattern inserts a condition to check that a
    collection index is within bound before it is used.
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre-fix and
post-fix version. If the count is different between pre- and post-fix
version, we consider the COB pattern found.

\vspace{1ex} The above patterns cover 8 out of 10 bug fixing patterns in
PAR. The other two patterns are (1) checking that an object is an instance of a
target class before casting it; and (2) replacing expression in an if condition
by another expression in the same scope. Pattern (1) cannot be analyzed because
Boa does not support the {\em instanceof} expression. Pattern (2)
requires us to track scope between pre-fix and post-fix version, which is not
easily supported by Boa.  Investigating 8 out of the
10 patterns should still provide a good idea on how common the patterns actually are.

\subsection{Results} \label{sec:freqfixpattern}

\begin{table}[!htb]
	\centering
	\begin{tabular}{lrr} 
		\hline
		& \textbf{\#Appearance} & \textbf{Percentage}\\
		\hline
		%Buggy Files & 46,301,429 \\ 
		AMP Pattern & 901,083 & 1.95\%\\
		MSM Pattern & 783,073 & 1.69\%\\
		COM Pattern & 270,128 & 0.58\%\\ 
		CBC Pattern & 1,959,377 & 4.23\%\\  
		IAO Pattern & 1,308,006 & 2.82\%\\  
		ANC Pattern & 1,340,561 & 2.90\%\\  
		AOB Pattern & 128,016 & 0.28\%\\  
		COB Pattern & 262,915 & 0.57\%\\   
		\hline
	\end{tabular}
	\caption{Frequency of Bug Fixing Patterns}\label{tab:freqpattern}

\end{table}

Table~\ref{tab:freqpattern} estimates the prevelance of PAR templates in the
Java dataset. The most common pattern we observed is CBC; and the least common
pattern is AOB. If we assume that these patterns never appear together, they
cover 14.78\% of buggy files. Kim et al. mentioned that the patterns that they
use cover almost 30\% of real patches~\cite{kim2013}. Although we only detect 8
out of 10 bug fixing patterns, it is unlikely the final two will cover the
difference. Moreover, our detection is imprecise, and likely represents an upper
bound;  This suggests that common bug fixing patterns might not generalize
beyond an initial dataset, and that more work remains to identify a sufficiently
expressive set of mutations for high-quality automated repair of Java.

\section{Statement-level mutations}
\begin{table*}
	\centering
	\resizebox{\textwidth}{100pt}{
		\begin{tabular}{|l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|} \hline
			&Assert&Break&Continue&Do& For&If&Label&Return&Case&Switch& Synch&Throw&Try&TypeDecl&While
			\\ \hline

\texttt{Assert}	&-&7.48&3.76&0.53&8.30&23.05&0.31&20.04&4.90&4.62&1.30&13.50&7.23&0.03&4.95
\\ \hline
\texttt{Break} &1.00&-&4.08&0.60&9.93&26.03&0.13&25.39&2.48&1.57&1.79&8.39&11.73&0.10&6.77
\\ \hline
\texttt{Continue} &1.74&9.42&-&1.28&11.39&18.25&0.35&22.60&3.80&2.85&2.17&8.98&9.42&0.11&7.63
\\   \hline
\texttt{Do} &0.81&5.26&6.60&0.00&9.44&14.21&0.18&15.86&3.73&1.67&1.97&5.88&6.39&0.03&27.98
\\ \hline
\texttt{For} &0.86&6.28&3.19&0.79&-&22.89&0.09&21.08&5.01&3.34&1.87&10.01&10.71&0.08&13.79
\\ \hline
\texttt{If}&1.64&8.43&2.87&0.60&13.49&-&0.24&26.46&7.45&4.80&2.85&9.89&15.11&0.08&6.11
\\ \hline
\texttt{Label} &1.30&8.33&7.86&1.11&5.18&22.85&-&15.17&3.05&2.04&14.62&10.45&4.16&0.09&3.79
\\ \hline
\texttt{Return} &1.13&9.41&3.11&0.49&13.33&27.24&0.24&-&5.59&3.65&2.55&14.91&12.61&0.12&5.61
\\ \hline
\texttt{Case} &0.78&2.84&2.84&0.39&10.27&31.79&0.16&22.40&-&0.46&2.07&7.37&11.69&0.08&6.87
\\ \hline
\texttt{Switch}&1.14&2.72&3.80&0.55&11.07&34.14&0.13&21.86&0.75&-&1.53&8.65&9.02&0.05&4.58
\\ \hline
\texttt{Synch} &0.80&6.57&2.28&0.43&10.21&24.18&0.05&19.77&6.35&2.07&-&9.16&12.16&0.04&5.93
\\ \hline
\texttt{Throw} &2.11&6.57&2.58&0.48&11.87&18.84&0.17&32.28&4.64&3.30&2.74&-&10.08&0.07&4.27
\\ \hline
\texttt{Try}&0.71&7.41&3.02&0.66&11.73&27.75&0.11&23.24&5.63&2.65&2.58&8.99&-&0.09&5.42
\\ \hline
\texttt{TypeDecl} &0.00&4.51&7.52&1.00&10.28&21.05&0.50&17.79&6.02&1.75&2.01&9.27&11.53&-&6.77
\\ \hline
\texttt{While}& 0.72&8.02&3.82&1.96&23.16&19.78&0.12&16.48&6.56&3.09&1.64&6.81&7.80&0.04&-
			\\ \hline
		\end{tabular}
		}
		\caption{Likeliness to replace the statement in the first column for the statement in the top row, in Java}\label{tab:likeliness}
\end{table*}

\begin{figure*}[!t] \centering \includegraphics[height=0.3\textheight,
	width=0.8\textwidth]{StmtType} \caption{For each kind of statement, how many
	bug fixing revisions add/delete statements of the same kind?}
	\label{fig:StmtType} \end{figure*}

Some automatic program repair approaches mutate source code using predefined
mutation operators such as addition, deletion and replacement. In order to
generate patches that make more sense to human developers, it would be
beneficial to analyze how developers mutate source code to fix bugs.

\vspace{1ex}\textbf{Approach.}
Because direct diffs are difficult to identify, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another across
this dataset.  To do this, for each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.
%   For each statement type (as tagged by the Boa infrastructure
% for Java), we process the results as follows: if number of occurrences of one
% statement type decreased post-fix and the number of another type increased, we
% say that the first statement type was replaced by the second statement type for
% that file. 
Note that this analysis doesn't distinguish the
replacement of the same statement kind, since we are counting the amount of
appearances of each statement kind.

We follow a similar approach to approximate a count of deletions and
insertions. 
For each bug fixing revision $r$ and each statement kind $k$, we count how many
statements of kind $k$ are there in revision $r$ and $r-1$ ($r-1$ means the
revision before revision $r$), and then compare the two numbers.

\vspace{1ex}
\noindent\textbf{Potential replacements.} Table~\ref{tab:likeliness} shows
the replacement likelihood for our dataset 
(each cell shows the percent of the time that the statement in the \emph{row}
was replaced by a statement of the type in the \emph{column}). For example, if
the cell in the DoStatement row (row 5) and AssertStatement column (column 2)
shows 0.81, meaning that Do statements were
replaced by Assert statements 0.81\% of the times.

Additional results (raw numbers not shown) show that the most common replacement
is replacing Return statement with If statement (appearing in 30,489 files). The
second most common replacement is replacing an If statement with a Return
statement, which had 28,536 appearances. 
By contrast, the least common replacement was an Assert
statement replacing a Type Declaration statement, which we did not observe. 
The second least common replacement was a Do statement for a Type
Declaration or a Label with a Type; we observed these once each.

The most common
\emph{replacer} statement kind (the statement kind that most commonly replaces
others) is the \texttt{If Statement} (101,366 appearances). The least
common \emph{replacer} is the \texttt{Type Declaration Statement} with 447
appearances.
The most common \emph{replacee} was the
\texttt{Return Statement} with 111,938 appearances; and the least common
\emph{replacee} was the \texttt{Type Declaration}, with 399 appearances.

\vspace{1ex}
\noindent \textbf{Potential deletions and insertions} From Figure~\ref{fig:StmtType}, we
can see that, expression, if, return, for and try catch statements are added
most often. Meanwhile, the same types of statements are also deleted most
often. This finding indicates that most bugs were fixed by changing the control
flow of the program.

\section{Discussion}

There exist threats to the validity of our study. Since we are using a DSL
provided by Boa to implement our analysis, the correctness of our analysis
depends on our programs as well as Boa itself. To mitigate the risk of errors in
implementation, we open source our implementation on
Github~\footnote{\url{https://github.com/}}.\todo{FIX THIS NOW.}  Some threats
to our study arise from limitations of Boa. For example, we rely on Boa to
identify bug fixing revisions; however, precisely accomplishing this is an open
problem. Another big limitation is that Boa does not provide an easy mechanism
to identify precise, statement-level diffs between revisions. Thus, our template
matching and analysis of code changes (by counting each statement kind) could be
inaccurate. However, our results should provide upper bounds.

We have characterized bug fixing changes in Java.
The findings of our study provide a set of useful insights for automatic
program repair tools in Java. It suggests that patterns proposed by the
state-of-the-art in generate and validate bug report for Java are not enough to
cover the entirety of bug fixes in Java. Adding more patterns is one solution,
but may not be a permanent one. Our finding on Java bug fixes characteristic
suggests that mutation based program repair may need to add new fields or
methods to fix bugs in Java and achieve a human-like written patch. A deeper
look suggests that it can also improve its search for a patch by having the knowledge of how
likely it is to replace a faulty statement by the different kinds of statements
available. Overall, our findings motivate additional study of the problem of
repair in Java, as assumptions that underlie approaches that target C are
unlikely to translate directly to this domain. 
% demonstrate that our knowledge of bug fixing
% properties in Java is still lacking and therefore more research on it are
% warranted.

%   We studied the applicability of the state-of-the-art in
% mutation operators to this dataset, 
% showing that 8 out of 10 bug fixing patterns in PAR cover only 15\% of bug
% fixes in this dataset.  Finally, investigated statement level mutations and
% summarized the likelihood of statement insertion, deletion, and replacement.
% These results suggest

% This study is directly applicable to automatic software repair. Most common
% tools for automatic software repair~\cite{kim2013,legoues2012} first localize a
% defect and then randomly select edits to 
% find a faulty statement using several different fault localization
% techniques~\cite{fry2010}, and once they have that faulty statement, they apply
% different kinds of edits to it, usually at random. Which makes it equally likely
% to replace an statement for another without looking at the kinds of statements
% being replaced. With this study, we are able to make a more accurate guess based
% on the phenomenon we observed from Table~\ref{tab:likeliness} and
% Figure~\ref{fig:StmtType}.

% For example, if we are fixing a Java program, and we have a faulty statement
% that happens to be a For Statement. Then we can look for the most likely
% statements to replace a For Statement (row 6): If statement (22.89\%
% likeliness), Return statement (21.08\%), and a While statement (13.79\% of
% likeliness to replace a For Statement). We also know that there is a 7.9\% of
% cases in which human programmers deleted this kind of statement, and in 10.6\%
% of the cases, a For statement was added to patch a bug. This way we provide
% guidelines to the automatic bug repair approaches to consider this on their
% effort to find a patch.

%\end{document}  % This is where a 'short' article might terminate

\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns

%\balancecolumns % GM June 2007
% That's all folks!
\end{document}